Model-free and Bayesian Ensembling Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL

### 2020-12-17 main.tex
- paper changes

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[
reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
amsmath,amssymb,amsfonts,clevref,
aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
]{revtex4-2}
\usepackage{wasysym}
\usepackage{graphicx, subcaption}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage[allcolors=blue]{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

	\title{Model-free and Bayesian Neural Network Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL}

@@

	\affiliation{IDA Lab, University of Salzburg, 5020 Salzburg, Austria}
	\author{Niky Bruchon}
	\affiliation{University of Trieste, Trieste, Italy}	
	\date{\today}% It is always \today, today,

@@

	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}.\\

@@

@@

	 Also, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\

@@

	\subsection{The physical set-up}

@@

	\subsection{The training environment}

@@

	and is limited by an $ |\ba_t|\leq \ba_\text{max}=1/12$.
	The training is done in episodes with a uniformly\footnote{in the normalised state-space} randomised initial state and a specific maximal length of taken steps afterwards - the horizon. For a reason discussed later, the horizon varies in the applications between either 10 or 500 steps. An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system. In our experiments, the maximal intensity was determined by manual expert optimisation. \\

@@

	\section{Deep Reinforcement learning}
	Most commonly, two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL). MFRL algorithms learn directly from the interaction with a system, while the MBRL algorithm learns a dynamics model from these interactions first. Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their applications is mainly restricted to simulated environments due to their low sample-efficiency. However, putting constraints on the algorithms, as discussed in \cref{s:Model-free reinforcement learning}, partly cures the sample inefficiency, turning them into practical tools for real-world problems.\\
	On the other hand, MBRL showing a much better sample efficiency. The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so-called model-bias problem \cite{Deisenroth2011}. In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample-efficiency and a low model-bias.\\
	In what follows, we denote the set of all states $\bs$ by $\states$ and the set of all actions $\ba$ by $\actions$. Relative to state and action vectors, $(\bs,\ba)$, we define a reward function $\reward : \states \times \actions \rightarrow \real$, a scalar $\discount \in (0,1]$ called discount factor and an initial state distribution $\initstate$. $\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$. The tuple \mbox{$\mdp = (\states,\actions,\transitions,\initstate,\reward,\discount)$} defines a Markov decision process.\\
	Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
	\begin{align}
		J(\policy^*)=\max J (\policy)  = 
		\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
	\end{align}
	where
	\begin{align}
		p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)
		\label{eq:trajectory_distribution}
	\end{align}

@@

	\subsection{Model free reinforcement learning}\label{s:Model-free reinforcement learning}

@@

	\subsection{Uncertainty aware DYNA-style reinforcement learning}\label{ss:Uncertainty aware DYNA-style reinforcement learning}
	%The original DYNA algorithm \cite{Kurutach2018} is modified here in several aspects.

@@

	In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$. In addition to the next state, the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change. In such a situation, the modelling can be simplified to model directly $\hat\transitions$}:

@@

	A schematic overview of our used method is shown in \cref{fig:MBRL_overview}. Initially, data is gathered by interaction with the real environment or read from a previous collection. An uncertainty aware model of the dynamics is trained, using 'anchored ensembling' \cite{Pearce2018} on the data (details provided in \cref{appendix:The impact of noise,ss:The uncertainty aware dynamics model}), which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account. It was motivated by the fact that pure epistemic ensemble techniques as \cite{Kurutach2018} were reported to be very sensitive to aleatoric noise \cite{Wang2019}. Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take these uncertainties into consideration using probabilistic ensembles. Our implementation is simpler while showing excellent performance in considered experiments.\\

@@

	After a defined number of training steps, called an epoch, the controller is validated on each model of the ensemble, individually. If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes. In this way, 'over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}. The training is stopped if the controller produces satisfying results on the real environment. Otherwise, a batch of new data is collected, and the model is improved, and consecutively a new epoch starts.

@@

	\subsection{Critical design decisions in MBRL}\label{ss:critical_design}

@@

	The 'anchored ensembling' method usually yields good results already with a small number of models. Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes}). The main goal is not to determine the exact posterior probability of the model, but, as already mentioned, providing the uncertainty to not over-train the controller at areas without sufficient data.\\

@@

	The number of data-points in a new batch to improve the model has to be chosen carefully. It is influenced by the number of randomly collected initial data-points (and then by the used policy $\policy$). The initialisation phase has to be selected not to small to minimise risks getting trapped in a local minimum for too long. Afterwards, from our experience, at least one full episode per epoch should be taken. We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in DYNA-style approaches. The maximal number of steps in our experimental runs per episode was ten. During the data acquisition on the real system, the latest learned policy $\policy$ was taken with some small additional Gaussian noise to improve exploration.

@@

	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment.

@@

	The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.
	\subsection{MFRL tests}

@@

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
		\caption{The training metrics of the \emph{AE-DYNA-SAC} on the FERMI FEL using a single network (dashed) and a double network (solid). The Bellman error (\cref{eq:minimize_bellmann_optimality}) and the state-value function (\cref{eq:state-value-function}) are shown. }
		\label{fig:NAF_convergence}
	\end{figure}
	In total, four tests were carried out, two using a single network and two using the double network architecture. 
	\Cref{fig:NAF_training} displays the results, averaged over the two tests \footnote{Considering the training is stochastic with random initial conditions, by chance, the training of the double network variant started above the threshold in both cases.}. A training of 100 episodes was accomplished. The number of iterations per episode is plotted, including the cumulative number of steps. In all episodes, during the training, the 95\% intensity reward threshold was surpassed, and therewith they were successfully finished. The evolution of the states is provided in \cref{fig:NAF_evolution_single,fig:NAF_evolution_double} for the single and double network. At the boundary of the domain, it takes time to learn, since many actions are mapped onto the same state\footnote{The states are clipped to $[0,1]$.}. The behaviour indicates that the double network converges faster and spends less time at the boundary.\\
	In the verification of 100 episodes, \cref{fig:NAF_verification}, both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance. Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps. The blue curves show the Bellmann error \cref{eq:minimize_bellmann_optimality}, which is comparable in both cases. The state-value function $V$ (\cref{eq:state-value-function}), which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value. In the single-network case, convergence is reached after around 1400 steps. In the plot, $V$ was averaged over 200 randomly selected states.
	\subsection{MBRL tests}
	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
		\caption{The training observables of the \emph{AE-DYNA-SAC} on the FERMI FEL.}
		\label{fig:AE-DYNA_observables}
	\end{figure}

@@

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
		\caption{The training observables of the \emph{ME-TRPO} on the FERMI FEL.}
		\label{fig:ME-TRPO_observables}
	\end{figure}

@@

	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented. Two different approaches were tested: a model-free and a model-based method.\\

@@

	Usage of the proposed methods in an operational way is attractive and could replace in the future the current optimisation method, which needs the destructive screen measurement. Hence an online retraining could be done and sacred time could be saved.\\ 

@@

	\begin{table}[h]%The best place to locate the table environment is directly after its first reference in text

@@

	The presented reinforcement learning methods hold tremendous promise for automatizing set-ups typical in accelerators. One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods.\\
	Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA-style} algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\

@@

	To provide the possibility for other laboratories to use the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.

	\section{Acknowledgements}
	We want to thank Giulio Gaio, who made these experiments possible and helped to set up and restore the FERMI FEL. Also, we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication. Finally, we want to thank the Land Salzburg for their financial support to create the context of the IDA lab, which made this work possible.

	\appendix
	\section{A non-linear standard control problem}
	To provide some transparency of these studies for other labs, we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.\\
	It is a non-linear low dimensional unsolved continuous control problem. Unsolved means there is no threshold for the reward to terminate an episode. The episode length, the horizon, is set to 200 steps. In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly concerning the noise handling. It is of importance when dealing with measurements on real systems.\\
	For statistical significance, all shown results were obtained using five different seeds. The average value and the standard deviation (shaded) are plotted. For this study, we assume that the problem is successfully solved if the cumulative reward surpasses a threshold of -200. A dashed green line indicates the threshold in the corresponding figures.
	%	  and a rolling average of 15 steps is used for better visualisation.

	\subsection{NAF2 details}\label{appendix:naf2}
	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network, as introduced in \cref{ss:Normalized advantage function} and in all other cases a single network is used. 
	The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
	\begin{equation}
		\ba(\bs) = \text{clip}\left(\mu_{\theta_{\text{targ}}}(\bs) + \text{clip}(\epsilon,-c,c), \ba_{Low}, \ba_{High}\right),
	\end{equation}
	where $\epsilon \sim \mathcal{N}(0, \sigma)$. $c>0$ denotes the clipping coefficient and $ \ba_{Low}, \ba_{High}$ the minimal and maximal possible action. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
	The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
	\begin{align}
		y(\reward_t,\bs_{t+1}, d_t) = \reward_t + \gamma (1 - d_t) \min_{i=1,2} V_{\theta_{i, \text{targ}}}(\bs_{t+1}),
	\end{align}
	with $d$ is 1 if the episode is finished and 0 otherwise.
	Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
	\begin{multline}
		L(\theta_i, {\mathcal D}) \\= \mathbb E_{(\bs_{t},\ba_{t},r_t,\bs_{t+1},d_t) \sim {\mathcal D}}{
			\Bigg( Q_{\theta_i}(\bs_{t},\ba_{t}) - y(r_t,\bs_{t+1},d_t) \Bigg)^2
		},
		\\
		i \in \{1,2\}
	\end{multline}
	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}(\bs)$.\\
	The results are shown in \cref{fig:comparsion_smoothing_small}. One sees the cumulative reward per episode for a training of a total of 100 episodes. As mentioned, the curve labelled \emph{clipping} corresponds to the double network, including \emph{smoothing} and shows the best overall stability during the training yielding a high reward quickly. Also, the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the slightly decreased stability. The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model-free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
	%	\begin{figure}[!h]
	%		\centering
	%		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
	%		\caption{Cumulative reward of different NAF implementations as discussed in the text.}
	%		\label{fig:comparsion_smoothing}
	%	\end{figure}
	\begin{figure}[!h]
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
		\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text on the \emph{inverted pendulum} without noise.}
		\label{fig:comparsion_smoothing_small}
	\end{figure}

@@

	\subsection{The impact of noise}\label{appendix:The impact of noise}

@@

	\subsubsection{Regression assuming homoskedastic Gaussian noise using `Anchored ensembling'}

@@

	new approaches to Bayesian inference. In the `Anchored ensembling' technique a regularisation

@@

		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the 'anchor ensembling'.}

>>>>

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[
reprint,nofootinbib,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
amsmath,amssymb,amsfonts,clevref,
aps,
%pra,
%prb,
%rmp,
prstab,
%prstper,
%floatfix,
]{revtex4-2}
\usepackage{wasysym}
\usepackage{graphicx, subcaption}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage[allcolors=blue]{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

	\title{Model-free and Bayesian Ensembling Model-based Deep Reinforcement Learning for Particle Accelerator Control Demonstrated on the FERMI FEL}

@@

	\affiliation{IDA Lab, University of Salzburg, 5020 Salzburg, Austria}
	\author{Niky Bruchon}\email{niky.bruchon@phd.units.it}
	\affiliation{University of Trieste, 34127 Trieste, Italy}	
	\date{\today}% It is always \today, today,

@@

	In this paper, we present the study carried out to solve the maximisation-problem of the radiation intensity generated by a seeded Free-Electron Laser (FEL) on the Free Electron laser Radiation for Multidisciplinary Investigations (FERMI) at \emph{Elettra Sincrotrone Trieste}. \Cref{fig:elletra_research} shows an aerial
	overview of the Elettra site and the FERMI free-electron laser buildings.\\

@@

		\includegraphics*[width=0.45\textwidth]{Figures/FERMI_AerialView}
		\caption{The Elettra research centre hosting the FERMI free electron laser \cite{Brochon2020}.}
		\label{fig:elletra_research}

@@

	 Thus, we show the operational applicability of purely model-based reinforcement learning in an accelerator control problem.\\

@@

	\subsection{The Physical Set-up}

@@

	\subsection{The Training Environment}

@@

	and is limited by an $ |\ba_t|\leq \ba_\text{max}=1/12$\footnote{The value was determined experimentally to guarantee reproducibility.}.
	The training is done in episodes with a uniformly\footnote{Sampled in the normalised state-space.} randomised initial state and a specific maximal length of taken steps afterwards - the horizon. For a reason discussed later, the horizon varies in the applications between either 10 or 500 steps. An early termination happens if a specific threshold is obtained, which lies in all cases at 95\% of the maximal achievable intensity of the system. In our experiments, the maximal intensity was determined through manual expert optimisation. \\

@@

	\section{Deep Reinforcement learning}
	Most commonly, two main categories of deep RL algorithms are distinguished: model-free RL (MFRL) and model-based RL (MBRL). MFRL algorithms learn directly from the interaction with a system, while MBRL algorithms learn a dynamics model from these interactions first. Despite the success of MFRL in several domains as robotics and video games \cite{Heess2017, Schulman2017,Silver2014,Lillicrap2015,OpenAI2018}, their applications is mainly restricted to simulated environments due to their low sample-efficiency. However, putting constraints on the algorithms, as discussed in \cref{s:Model-free reinforcement learning}, partly cures the sample inefficiency, turning them into practical tools for real-world problems.\\
	On the other hand, MBRL showing a much better sample efficiency. The challenge for these methods is to learn an accurate model, which turns out to be rather hard and leads to the so-called model-bias problem \cite{Deisenroth2011}. In \cref{ss:Uncertainty aware DYNA-style reinforcement learning} we discuss attempts to capture the model uncertainty to alleviate this obstacle by making use of model ensembles and Bayesian neural networks resulting in algorithms, exhibiting a high sample-efficiency and a low model-bias.\\
	In what follows, we denote the set of all states $\bs$ by $\states$ and the set of all actions $\ba$ by $\actions$. Relative to state and action vectors, $(\bs,\ba)$, we define a reward function $\reward : \states \times \actions \rightarrow \real$, a scalar $\discount \in (0,1]$ called discount factor and an initial state distribution $\initstate$. $\transitions(\bs_{t+1}|\bs_t,\ba_t)$ characterises the probability to end in a state $\bs_{t+1}$ if an action $\ba_t$ is taken at state $\bs_t$. The tuple \mbox{$\mdp = (\states,\actions,\transitions,\initstate,\reward,\discount)$} defines a Markov decision process.\\
	Goal of reinforcement learning is to find a $\policy^*: \bs\mapsto \ba$, which is the solution of:
	\begin{align}
		J(\policy^*)=\max J (\policy)  = 
		\mathbb E_{\traj \sim p_{\policy}}\left[\sum_{t=0}^{\horizon}\discount^t r(\bs_t,\ba_t)\right],\label{eq:cumulative_reward}
	\end{align}
	where
	\begin{align}
		p_\policy = \initstate(\bs_0)\prod_{t_0}^\horizon\policy(\ba_t,\bs_t)\transitions(\bs_{t+1}|\bs_t,\ba_t)
		\label{eq:trajectory_distribution}
	\end{align}

@@

	\subsection{Model-free Reinforcement Learning}\label{s:Model-free reinforcement learning}

@@

	\subsection{Uncertainty Aware DYNA-style Reinforcement Learning}\label{ss:Uncertainty aware DYNA-style reinforcement learning}
	%The original DYNA algorithm \cite{Kurutach2018} is modified here in several aspects.

@@

	In MFRL the dynamics $\transitions$ of an environment is learned implicitly, while in model-based reinforcement learning MBRL $\transitions$ is learned explicitly and approximated as $\hat \transitions$. In addition to the next state, the reward is included in the model \footnote{In some applications $\reward$ can be deduced from the current state or the state change. In such a situation, the modelling can be simplified to model directly $\hat\transitions$.}:

@@

	A schematic overview of our used method is shown in \cref{fig:MBRL_overview}. Initially, data is gathered by interaction with the real environment or read from a previous collection. An uncertainty aware model of the dynamics is trained, using `anchored ensembling' \cite{Pearce2018} on the data (details provided in \cref{appendix:The impact of noise,ss:The uncertainty aware dynamics model}), which allows taking the aleatoric (measurement errors) as well as the epistemic uncertainty (lack of data) into account. It was motivated by the fact that pure epistemic ensemble techniques as \cite{Kurutach2018} were reported to be very sensitive to aleatoric noise \cite{Wang2019}. Alternatives as \cite{Chua2018,Janner2019,Wang2019a} also take these uncertainties into consideration using probabilistic ensembles. Our implementation is simpler while showing excellent performance in considered experiments.\\

@@

	After a defined number of training steps, called an epoch, the controller is validated on each model of the ensemble, individually. If there is no improvement in a number smaller than the total number of the models, the controller might be tested on the real environment for several episodes. In this way, `over-fitting' on a wrong model is avoided, as discussed in \cite{Kurutach2018}. The training is stopped if the controller produces satisfying results on the real environment. Otherwise, a batch of new data is collected, and the model is improved, and consecutively a new epoch starts.

@@

	\subsection{Critical Design Decisions in MBRL}\label{ss:critical_design}

@@

	The `anchored ensembling' method usually yields good results already with a small number of models. Empirical results showed that already three models were satisfactory to see definite improvements over a single network approach (see \cref{fig:Compare_models_sizes}). The main goal is not to determine the exact posterior probability of the model, but, as already mentioned, providing the uncertainty to not over-train the controller at areas without sufficient data.\\

@@

	The number of data-points in a new batch to improve the model has to be chosen carefully. It is influenced by the number of randomly collected initial data-points (and then by the used policy $\policy$). The initialisation phase has to be selected not to small to minimise risks getting trapped in a local minimum for too long. Afterwards, from our experience, at least one full episode per epoch should be taken. We decided to use a short horizon to diminish the impact of the compound error \cite{Janner2019}, the accumulation error following a wrong model, as well known in \emph{DYNA}-style approaches. The maximal number of steps in our experimental runs per episode was ten. During the data acquisition on the real system, the latest learned policy $\policy$ was taken with some small additional Gaussian noise to improve exploration.

@@

	\emph{ME-TRPO} and \emph{AE-DYNA} are tested and compared on the real environment. The \emph{NAF2} algorithm, the representative for highly sample efficient MFRL algorithms, is discussed first.

@@

	\subsection{MFRL Tests}

@@

		\includegraphics*[width=0.5\textwidth]{Figures/FERMI_all_experiments_NAF_convergence.pdf}
		\caption{The training metrics of the \emph{AE-DYNA-SAC} on the FERMI FEL using a single network (dashed) and a double network (solid). The Bellman error (\cref{eq:minimize_bellmann_optimality}) and the state-value function (\cref{eq:state-value-function}) are shown. }
		\label{fig:NAF_convergence}
	In total, four tests were carried out, two using a single network and two using the double network architecture. 
	\Cref{fig:NAF_training} displays the results, averaged over the two tests \footnote{Considering the training is stochastic with random initial conditions, by chance, the training of the double network variant started above the threshold in both cases.}. A training of 100 episodes was accomplished. The number of iterations per episode is plotted, including the cumulative number of steps. In all episodes, during the training, the 95\% intensity reward threshold was surpassed, and therewith they were successfully finished. The evolution of the states is provided in \cref{fig:NAF_evolution_single,fig:NAF_evolution_double} for the single and double network. At the boundary of the domain, it takes time to learn, since many actions are mapped onto the same state\footnote{The states are clipped to $[0,1]$.}. The behaviour indicates that the double network converges faster and spends less time at the boundary.\\
	In the verification of 100 episodes, \cref{fig:NAF_verification}, both algorithms show similar performance, while the double network needed less training steps, and reveals a more stable overall performance. Additionally, the convergence metrics of the two algorithms is plotted in \cref{fig:NAF_convergence} against the number of training steps. The blue curves show the Bellmann error \cref{eq:minimize_bellmann_optimality}, which is comparable in both cases. The state-value function $V$ (\cref{eq:state-value-function}), which is a direct output of the neural net (\cref{eq:state-action-value-approxiation}), converges to a reasonable value for the double network within the shown 700 steps, whereas the single network seems to overestimate the value. In the single-network case, convergence is reached after around 1400 steps. In the plot, $V$ was averaged over 200 randomly selected states.
	\subsection{MBRL Tests}
		\includegraphics*[width=0.5\textwidth]{Figures/AE-DYNA_observables.pdf}
		\caption{The training observables of the \emph{AE-DYNA-SAC} on the FERMI FEL. Detail are provided in the text.}
		\label{fig:AE-DYNA_observables}

@@

	\begin{figure}
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/ME-TRPO_observables.pdf}
		\caption{The training observables of the \emph{ME-TRPO} on the FERMI FEL. Detail are provided in the text.}
		\label{fig:ME-TRPO_observables}
	\end{figure}

@@

	In this paper, the applicability of deep reinforcement learning on the optimisation of the FERMI FEL intensity was presented. Two different approaches were tested: model-free and model-based.\\

@@

	Usage of the proposed methods in an operational way is attractive and could replace in the future the current optimisation method, which needs the destructive screen measurement. Hence an online retraining could be done and valuable time could be saved.\\ 

@@

	\begin{table}[!h]%The best place to locate the table environment is directly after its first reference in text

@@

	The presented reinforcement learning methods hold tremendous promise for automatizing set-ups typical in accelerators. One objective of this work was to provide some suggestions on how to make advanced deep reinforcement learning techniques work on a real set-up by adapting available methods. This was demonstrated on the FERMI FEL intensity optimization problem.\\
	Regularly, control problems in accelerators have short horizons, which makes \emph{DYNA}-style algorithms as the \emph{ME-TRPO} or our \emph{AE-DYNA} with their high representational power and excellent sample efficiency appealing choices.\\

@@

	To provide the possibility for other laboratories to profit form the stated methods, the code was released in \url{https://github.com/MathPhysSim/FERMI_RL_Paper} and \cite{Hirlaender2020b}.

	\section{Acknowledgements}
	We want to thank Giulio Gaio, who made these experiments possible and helped to set up and restore the FERMI FEL. Also, we are thankful for the valuable feedback from Alexander Scheinker on the draft of this publication. Finally, we want to thank the Land Salzburg for their financial support to create the IDA Lab, which made this work possible.

	\appendix
	\section{A Non-linear Standard Control Problem}
	To provide some transparency of these studies for other labs, we provide results on a famous classical standard control problem \cite{Furutaa}, the \emph{inverted pendulum}.\\
	It is a non-linear low dimensional unsolved continuous control problem. Unsolved means there is no threshold for the reward to terminate an episode. The episode length, the horizon, is set to 200 steps. In the following several tests were carried out on the \emph{inverted pendulum} to demonstrate the improvements of the selected algorithms, mainly concerning the noise handling. It is of importance when dealing with measurements on real systems.\\
	For statistical significance, all shown results were obtained using five different seeds. The average value and the standard deviation (shaded) are plotted. For this study, we assume that the problem is successfully solved if the cumulative reward surpasses a threshold of -200. A dashed green line indicates the threshold in the corresponding figures.
	%	  and a rolling average of 15 steps is used for better visualisation.

	\subsection{NAF2 Details}\label{appendix:naf2}
	%We add artificial Gaussian noise $\gauss(0,0.05)$ on the state.
	We compare the different \emph{NAF} variants: \emph{Clipping}, \emph{No-clipping-smoothing}, \emph{No-clipping-no-smoothing}, where \emph{clipping} indicates the use of the double network, as introduced in \cref{ss:Normalized advantage function} and in all other cases a single network is used. 
	The term \emph{smoothing} indicates that a small clipped noise is added on the actions to stabilize the network training as:
	\begin{equation}
		\ba(\bs) = \text{clip}\left(\mu_{\theta_{\text{targ}}}(\bs) + \text{clip}(\epsilon,-c,c), \ba_{Low}, \ba_{High}\right),
	\end{equation}
	where $\epsilon \sim \mathcal{N}(0, \sigma)$. $c>0$ denotes the clipping coefficient and $ \ba_{Low}, \ba_{High}$ the minimal and maximal possible action. This method was used already in \cite{fujimoto2018addressing} to improve the deterministic policy gradient \cite{Silver2014}.
	The double network was used in \cite{fujimoto2018addressing,Haarnoja2018a} and is done in the following way:
	\begin{align}
		y(\reward_t,\bs_{t+1}, d_t) = \reward_t + \gamma (1 - d_t) \min_{i=1,2} V_{\theta_{i, \text{targ}}}(\bs_{t+1}),
	\end{align}
	with $d$ is 1 if the episode is finished and 0 otherwise.
	Then both are learned by regressing to this target (using the tuples from the data buffer $\mathcal D$ ):
	\begin{multline}
		L(\theta_i, {\mathcal D}) \\= \mathbb E_{(\bs_{t},\ba_{t},r_t,\bs_{t+1},d_t) \sim {\mathcal D}}{
			\Bigg( Q_{\theta_i}(\bs_{t},\ba_{t}) - y(r_t,\bs_{t+1},d_t) \Bigg)^2
		},
		\\
		i \in \{1,2\}
	\end{multline}
	and the policy is obtained via $\max_\ba Q_{\theta_1}(\bs,\ba) = \mu_{\theta_1}(\bs)$.\\
	The results are shown in \cref{fig:comparsion_smoothing_small}. One sees the cumulative reward per episode for a training of a total of 100 episodes. As mentioned, the curve labelled \emph{clipping} corresponds to the double network, including \emph{smoothing} and shows the best overall stability during the training yielding a high reward quickly. Also, the \emph{smoothed} single network, labelled \emph{No-clipping-smoothing}, shows good and comparable performance, except for the slightly decreased stability. The worst performance is achieved without smoothing and a single network (\emph{No-clipping-no-smoothing}), nevertheless the result is competing with state of the art model-free methods as \cite{BarthMaron2018} as the benchmark in the \emph{leaderboard} of openai gym \cite{Brockman2016}.
	%	\begin{figure}[!h]
	%		\centering
	%		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_smoothing}
	%		\caption{Cumulative reward of different NAF implementations as discussed in the text.}
	%		\label{fig:comparsion_smoothing}
	%	\end{figure}
	\begin{figure}[!h]
		\centering
		\includegraphics*[width=0.5\textwidth]{Figures/Comparison_naf}
		\caption{Cumulative reward of different \emph{NAF} implementations as discussed in the text on the \emph{inverted pendulum} without noise.}
		\label{fig:comparsion_smoothing_small}
	\end{figure}

@@

	\subsection{The Impact of Noise}\label{appendix:The impact of noise}

@@

	\subsubsection{Regression assuming homoskedastic Gaussian noise using `anchored ensembling'}

@@

	new approaches to Bayesian inference. In the `anchored ensembling' technique a regularisation

@@

		\caption{Cumulative reward of \emph{AE-DYNA-SAC} on the \emph{inverted pendulum} with artificial noise using the `anchor ensembling'.}
